\HeaderA{mice.impute.logreg2}{Elementary Imputation Method: Logistic Regression}{mice.impute.logreg2}
\keyword{misc}{mice.impute.logreg2}
\begin{Description}\relax
Imputes univariate missing data using logistic regression.
\end{Description}
\begin{Usage}
\begin{verbatim}
    imp <- mice.impute.logreg2(y, ry, x)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{y}] Incomplete data vector of length n
\item[\code{ry}] Vector of missing data pattern of length n (FALSE=missing,
TRUE=observed)
\item[\code{x}] Matrix (n x p) of complete covariates.

\end{ldescription}
\end{Arguments}
\begin{Details}\relax
Imputation for binary response variables by the Bayesian 
logistic regression model. See Rubin (1987, p. 169-170) for
a description of the method.
The method consists of the following steps:
\Enumerate{
\item Fit a logit, and find (bhat, V(bhat))
\item Draw BETA from N(bhat, V(bhat))
\item Compute predicted scores for m.d., i.e. logit-1(X BETA)
\item Compare the score to a random (0,1) deviate, and mice.impute.
}
This method uses direct minimization of the likelihood function
by means of V\&R function logitreg (V\&R, 2nd ed, p. 293).
\end{Details}
\begin{Value}
\begin{ldescription}
\item[\code{imp}] A vector of length nmis with imputations (0 or 1).
\end{ldescription}
\end{Value}
\begin{Note}\relax
An alternative is mice.impute.logreg.
\end{Note}
\begin{Author}\relax
Stef van Buuren, Karin Oudshoorn, 2000
\end{Author}
\begin{References}\relax
Van Buuren, S. \& Oudshoorn, C.G.M. (2000). Multivariate Imputation by Chained Equations: 
MICE V1.0 User's manual. Report PG/VGZ/00.038, TNO Prevention and Health, Leiden.

Brand, J.P.L. (1999). Development, Implementation and Evaluation
of Multiple
Imputation Strategies for the Statistical Analysis of Incomplete
Data Sets.
Ph.D. Thesis, TNO Prevention and Health/Erasmus University Rotterdam. ISBN 90-74479-08-1. 

Venables, W.N. \& Ripley, B.D. (1997). Modern applied statistics with S-Plus (2nd ed). Springer, Berlin.
\end{References}
\begin{SeeAlso}\relax
\code{\LinkA{mice}{mice}}, \code{\LinkA{logitreg}{logitreg}}, \code{\LinkA{mice.impute.logreg}{mice.impute.logreg}}
\end{SeeAlso}

