\HeaderA{mice.impute.norm}{Elementary Imputation Method: Linear Regression Analysis}{mice.impute.norm}
\keyword{misc}{mice.impute.norm}
\begin{Description}\relax
Imputes univariate missing data using linear regression analysis
\end{Description}
\begin{Usage}
\begin{verbatim}
mice.impute.norm(y, ry, x)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{y}] Incomplete data vector of length n
\item[\code{ry}] Vector of missing data pattern (FALSE=missing, TRUE=observed)
\item[\code{x}] Matrix (n x p) of complete covariates.
\end{ldescription}
\end{Arguments}
\begin{Details}\relax
Draws values of beta and sigma for Bayesian linear regression imputation 
of y given x according to Rubin p. 167.
\end{Details}
\begin{Value}
A vector of length nmis with imputations.
\end{Value}
\begin{Note}\relax
Using mice.impute.norm for all columns gives results similar to Schafer's norm 
method (Schafer, 1997), though much slower.
\end{Note}
\begin{Author}\relax
Stef van Buuren, Karin Oudshoorn, 2000
\end{Author}
\begin{References}\relax
Van Buuren, S. \& Oudshoorn, C.G.M. (2000). Multivariate Imputation by Chained Equations: 
MICE V1.0 User's manual. Report PG/VGZ/00.038, TNO Prevention and Health, Leiden.
Brand, J.P.L. (1999). Development, Implementation and Evaluation of Multiple Imputation Strategies for the Statistical Analysis of Incomplete Data Sets. Ph.D. Thesis, TNO Prevention and Health/Erasmus University Rotterdam. ISBN 90-74479-08-1. 

Schafer, J.L. (1997). Analysis of incomplete multivariate data. London: Chapman \& Hall.
\end{References}

