\HeaderA{mice.impute.lda}{Elementary Imputation Method: Linear Discriminant Analysis}{mice.impute.lda}
\keyword{misc}{mice.impute.lda}
\begin{Description}\relax
Imputes univariate missing data using linear discriminant analysis
\end{Description}
\begin{Usage}
\begin{verbatim}
mice.impute.lda(y, ry, x)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{y}] Incomplete data vector of length n
\item[\code{ry}] Vector of missing data pattern (FALSE=missing, TRUE=observed)
\item[\code{x}] Matrix (n x p) of complete covariates.
\end{ldescription}
\end{Arguments}
\begin{Details}\relax
Imputation of categorical response variables by linear discriminant
analysis. This function uses the Venables/Ripley functions
lda and predict.lda to compute posterior probabilities for
each incomplete case, and draws the imputations from this 
posterior.
\end{Details}
\begin{Value}
A vector of length nmis with imputations.
\end{Value}
\begin{Section}{Warning}
The function does not incorporate the variability of the discriminant 
weight, so it is not 'proper' in the sense of Rubin. For small samples
and rare categories in the y, variability of the mice.imputed data could 
therefore be somewhat underestimated.
\end{Section}
\begin{Note}\relax
This function can be called from within the Gibbs sampler by specifying 
'lda' in the imputationMethod argument. 
This method is usually faster and uses less resources than 
\code{\LinkA{mice.impute.polyreg}{mice.impute.polyreg}}.
\end{Note}
\begin{Author}\relax
Stef van Buuren, Karin Oudshoorn, 2000
\end{Author}
\begin{References}\relax
Van Buuren, S. \& Oudshoorn, C.G.M. (2000). Multivariate Imputation by Chained Equations: 
MICE V1.0 User's manual. Report PG/VGZ/00.038, TNO Prevention and Health, Leiden.

Brand, J.P.L. (1999). Development, Implementation and Evaluation of
Multiple Imputation Strategies for the Statistical Analysis of
Incomplete Data Sets.
Ph.D. Thesis, TNO Prevention and Health/Erasmus University Rotterdam. ISBN 90-74479-08-1. 

Venables, W.N. \& Ripley, B.D. (1999). Modern applied statistics with S-Plus (3rd ed). Springer, Berlin.
\end{References}
\begin{SeeAlso}\relax
\code{\LinkA{mice}{mice}}, \code{\LinkA{lda}{lda}}, \code{\LinkA{predict.lda}{predict.lda}}
\end{SeeAlso}

